Echoes from the scattering of wavepackets on wormholes

### 2018-12-16 paper.tex
no message

After fixing the values of the width to be $\sigma=0.9185r_g$, the initial position of the Gaussian at $x_0=60.0r_g$, $r_0=20.0r_g$ and the initial conditions in \eqref{eq:scalar_init_cond}, we show the time-dependent solution of \eqref{eq:wave_scalar} in Fig.~\ref{fig:bh_sol}, where we distinguish the incident, transmitted and reflected parts of the solution. It is important to observe the absence of spurious late time reflections and interferences due to the implementation of perfectly matching layers (PMLs) in the outermost regions of our simulation box (see the details of our setup for PMLs in \citep{Frolov:2017asg}). We observe the main features of the reflected signal in Fig.~\ref{fig:asympt_sol}, where the asymptotic behavior of the signal shows a sharp transient is a consequence of the collision against the potential wall, and the ringing of quasinormal modes occuring right after the reflection in agreement with \citep{Petrich:1985csm}.

@@

The amplitude of the outgoing signals depends on the width of the incident Gaussian wavepacket, this is visible in Fig.~\ref{fig:asympt_two_sigmas} where we plot the asymptotic solutions for two different ingoing wavepackets: one with $\sigma=0.6495r_g$ in the left panel and a second one with $\sigma=5.196r_g$ in the right. After subtracting the outgoing solution for a black hole, the presence of a train of wavelets, widely known as echoes, is very clear. The signal is not a straight line after the transient, as we can observe in the right panel of this figure. Therefore, subtracting the black hole outgoing pulse (i.e., the case in which there is only one potential wall) turns to be a convenient way to clean the signal from back reflections due to the ``tails'' of the potential barrier, especially since in this case the echoes are four orders of magnitude smaller than the transient.

As shown in subsection \ref{subsec:bh}, the values of reflectivity and transmissivity determine the frequencies going through the walls of the potential cavity: most of the power of an incident pulse with large $\sigma$ is in the low frequency domain, and therefore it will be reflected. Moreover, the cavity is transparent to high frequency signals, which are dominant in the pulses with small $\sigma$. 
\begin{figure*}
\centering
\subfigure{
\includegraphics[width=.45\textwidth]{figures/Scalar_echo_w_06495.pdf}} \,
\subfigure{
\includegraphics[width=.45\textwidth]{figures/Scalar_echo.pdf}}
\caption{\label{fig:asympt_two_sigmas} Left panel: Asymptotic evolution of the quadrupole mode $\mathcal{U}_{20}$ for $\sigma=0.6495r_g$. Echoes are plotted in the upper corner of the plot, showing them along with their Hilbert envelope (in green) and maximum amplitudes (in red). Right panel: Asymptotic evolution of the quadrupole mode $\mathcal{U}_{20}$ for $\sigma=5.196r_g$. In contrast with the left panel, echoes are four orders of magnitude smaller than the transient.} 
\end{figure*}
In either of these extremal scenarios, QNMs cannot be sourced by internal reflections and thus, the amplitude of the echoes is not generically large. Now we proceed to reconstruct the asymptotic spectrum using the reflectivity and transmissivity defined in \eqref{eq:ref_and_trans} and the compare it with the Fourier transformed asymptotic solutions of \eqref{eq:wave_scalar_wh}.

>>>>

After fixing the values of the width to be $\sigma=0.9185r_g$, the initial position of the Gaussian at $x_0=60.0r_g$, $r_0=20.0r_g$ and the initial conditions in \eqref{eq:scalar_init_cond}, we show the time-dependent solution of \eqref{eq:wave_scalar} in Fig.~\ref{fig:bh_sol}, where we distinguish the incident, transmitted and reflected parts of the solution. It is important to observe the absence of spurious late time reflections and interferences due to the implementation of perfectly matching layers (PMLs) in the outermost regions of our simulation box (see the details of our setup for PMLs in \citep{Frolov:2017asg}). We observe the main features of the reflected signal in Fig.~\ref{fig:asympt_sol}, where the asymptotic behavior of the signal shows a sharp transient is a consequence of the collision against the potential wall, and the ringing of quasinormal modes occurring right after the reflection in agreement with \citep{Petrich:1985csm}.

@@

The scalar flux is shown in the right panel, we observe that the only source of scalar radiation comes from the first collision of the Gaussian wavepacket against the barrier in the right hand side (the ingoing flux is colored in black at the bottom of the contour plot). A sequence of reflections occurs within the potential barriers, which decay in intensity with time as the cavity leaks energy to the exterior.

\begin{figure*}
\centering
\subfigure{
\includegraphics[width=.45\textwidth]{figures/Scalar_echo_w_06495.pdf}} \,
\subfigure{
\includegraphics[width=.45\textwidth]{figures/Scalar_echo.pdf}}
\caption{\label{fig:asympt_two_sigmas} Left panel: Asymptotic evolution of the quadrupole mode $\mathcal{U}_{20}$ for $\sigma=0.6495r_g$. Echoes are plotted in the upper corner of the plot, showing them along with their Hilbert envelope (the curves in green) and maximum amplitudes (in red). Right panel: Asymptotic evolution of the quadrupole mode $\mathcal{U}_{20}$ for $\sigma=5.196r_g$. In contrast with the left panel, echoes are four orders of magnitude smaller than the transient. The amplitude of the transient in the right figure has decreased with respect to the one in the left panel.} 
\end{figure*}

The amplitude of the outgoing signals depends on the width of the incident Gaussian wavepacket, this is visible in Fig.~\ref{fig:asympt_two_sigmas} where we plot the asymptotic solutions for two different ingoing wavepackets: one with $\sigma=0.6495r_g$ in the left panel and a second one with $\sigma=5.196r_g$ in the right. After subtracting the outgoing solution for a black hole, the presence of a train of wavelets, widely known as echoes, is very clear. The signal is not a straight line after the transient, as we can observe in the right panel of this figure. Therefore, subtracting the black hole outgoing pulse (i.e., the case in which there is only one potential wall) turns to be a convenient way to clean the signal from back reflections due to the ``tails'' of the potential barrier, especially since in this case the echoes are four orders of magnitude smaller than the transient. In both panels, we plot the variable $\Delta\mathcal{U}_{20}|_{\mathrm{clean}}\equiv\mathcal{U}_{20}|_{\mathrm{original}}-\mathcal{U}^{\mathrm{bh}}_{20}$ in the right upper corner of the figures to represent the echoes and their net amplitude without backscattering effects.

As shown in subsection \ref{subsec:bh}, the curves of reflectivity and transmissivity determine the frequencies going through the barriers of the potential cavity: most of the power of an incident pulse with large $\sigma$ is in the low frequency domain, and therefore it will be reflected. Moreover, the cavity is transparent to high frequency signals, which are dominant in the pulses with small $\sigma$. In either of these extremal scenarios, QNMs cannot be sourced by internal reflections and thus, the amplitude of the echoes is not large in general.\\

\begin{figure}[t!]
\centering
\includegraphics[width=.45\textwidth]{figures/Reconst_omega_scalar.pdf}
\caption{\label{fig:rec_scalar} Comparing the reconstruction of the spectrum by the geometrical optics approximation in \eqref{eq:reconst} (in red) with the Fourier transform of the full asymptotic solution $\tilde{\mathcal{U}}_{20}^{\mathrm{echo}}$ (in blue). The color stripes indicate a few of the first QNM frequency peaks corresponding to $\omega_3=0.251M^{-1}$ in golden rod, $\omega_4=0.207M^{-1}$ in peru, $\omega_5=0.169M^{-1}$ in plum, $\omega_6=0.133M^{-1}$ in light blue and $\omega_7=0.092M^{-1}$ in rosy brown, these are the real parts of the QNM frequencies for $n=3,4,5,6~\&~7$ calculated in \citep{PhysRevD.46.4179}.}
\end{figure}

Now we proceed to reconstruct the asymptotic spectrum by following the geometrical optics relation in \eqref{eq:reconst} considering the reflectivity and transmissivity operators defined in \eqref{eq:ref_and_trans}. Then, the outcome should be compared with the spectral content of the asymptotic wave solutions of \eqref{eq:wave_scalar_wh}. We calculate the Fourier transforms of both the Gaussian incident pulse $\Phi_0(\omega)=\mathcal{F}[\exp\left((x-x_0)^2/2\sigma^2\right)]$ for $\sigma=0.6495r_g$ and $x_0=60.0r_g$, and the asymptotic solution $\tilde{\mathcal{U}}_{20}^{\mathrm{echo}}\equiv\mathcal{F}[\mathcal{U}_{20}(t,+\infty)]$, which includes the echoes.\\

After applying the reconstruction expression in \eqref{eq:reconst} up to $i=0$ (in red), and comparing the outcome with $\tilde{\mathcal{U}}_{20}^{\mathrm{echo}}$, we show the reconstructed spectrum and $\tilde{\mathcal{U}}_{20}^{\mathrm{echo}}$ (in blue) in Fig.~\ref{fig:rec_scalar}. The low frequency oscillation peaks in the blue spectrum correspond to the finite size of the simulation box and should be ignored. Notice that the spectrum reconstructed employing the geometrical optics approximation gives the overall shape of the spectrum with decent precision but not the QNM frequency peaks, these appear in the same frequency interval where the reflectivity and transmissivity curves intersect in Fig.~\ref{fig:RT_scalar}. It is, therefore, reasonable to talk about a ``sweet spot'' in the frequency domain where the cavity maximizes the amplitude of the echoes. Intuitively, after observing the results in Fig.~\ref{fig:asympt_two_sigmas}, it is possible to identify a similar ``sweet spot'' in the parameter space space for the widths of the incident Gaussian wavepackets. However, it is also necessary to not only compare the amplitude of each individual echo with $\sigma$, but also the ratio between the amplitude of the echo with the amplitude of the transient for each value of the width.
